Exploring Conformational Dynamics and Related Time Scales via Nonequilibrium Molecular Dynamics Simulation

### 2013-08-16 writedown/npt-neq/alanine.tex
backup

The dynamical nonequilibrium molecular dynamics simulation (D-NEMD)
was firstly developed by G. Ciccotti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought}, and
recently successfully applied to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. D-NEMD provides a way of evaluating
nonequilibrium observables by doing an ensemble of nonequilibrium MD
simulations.  It can be used for analyzing the
conformational changing under a generally time-dependent EF.
In the present paper, we study the minimal model peptide
segment -- alanine dipeptide. We take the approach of starting from a
single peptide piece rather than a more realistic chain, firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding all possible interplay between
different peptide segments along a chain.
Secondly, the current computational resource available is not enough
to study peptide chains by D-NEMD, since one needs to simulate an
ensemble of MD trajectories (typically a few thousands), so it
requires much higher computational cost than producing a single
equilibrium MD trajectory.

@@

@@

The Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment of the alanine dipeptide
reaches 85\% of the maximum value in only 20~ps, which is comparable to
the warm-up time $t_{init}$, and is 25 times
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to the 
maximum value, that reads
6.8~Debye, and is  corresponding to a dipole energy of ca.
$-13.6$~kJ/mol. The main free energy barriers among the
metastable states are reported to be of order $20\sim 40$~kJ/mol~\cite{bohner2012algorithm},
which is comparable to the dipole energy. This explains the 
conformational changes under the electric field.
We calculate the averaged dipole moment of different metastable state
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
Under a constant EF, the system will be driven towards those
conformations that with higher dipole moment, because the
energy of the system will be lowered by aligning the the dipole
to the external EF.
It is clear that
the $\alpha$-helix and left-handed $\alpha$-helix are of higher
dipole moments than $\beta$-sheet conformations, so the probability
of being in the $\alpha$-helix conformations were
observed being increased under the EF.  $A_1$ and $C$ are the most
probable conformations because their dipole moments are the highest
among all metastable conformation. Moreover, the $B_2$  vanishes, because
its dipole is significantly lower than other conformations.
% This also explains the two-time scale behavior of the molecular dipole
% moment.
From 0 to 20~ps, the probability of
lowest dipole conformation $B_2$ quickly
vanishes, so there is a sharp increment of the average molecular
dipole moment. Then until 500~ps, since the conformation slowly migrates from
$B_1$ to $A_1$ and $C$, and also since the dipole moment $A_1$ and $C$
is higher than $B_1$, the average molecular dipole moment slowly increases
to the maximum value.

>>>>

The dynamical nonequilibrium molecular dynamics simulation (D-NEMD)
was firstly developed by G. Ciccotti and his
colleagues~\cite{ciccotti1975direct, ciccotti1979thought}, and
recently successfully applied to the study of
hydrodynamics~\cite{orlandini2011hydrodynamics,
  orlandini2011hydrodynamics-01}. D-NEMD provides a way of evaluating
nonequilibrium observables by doing an ensemble of nonequilibrium MD
simulations.  It can be used for analyzing the
conformational changing under a generally time-dependent EF.
In the present paper, we study a small model peptide
segment -- alanine dipeptide. We take the approach of starting from a
single peptide piece rather than a more realistic chain, firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding all possible interplay between
different peptide segments along a chain.
Secondly, the current computational resource available is not enough
to study peptide chains by D-NEMD, since one needs to simulate an
ensemble of MD trajectories (typically a few thousands), so it
requires much higher computational cost than producing a single
equilibrium MD trajectory.

@@

\begin{table}
  \centering
  \begin{tabular*}{0.4\textwidth}{@{\extracolsep{\fill}}lc}\hline\hline
    Direction        & time scale [ ps ] \\\hline
    $A_1\rightarrow A_2$        &       $0\sim 10$      \\
    $B_1\rightarrow B_2$        &       $0\sim 10$      \\    
    $A_2\rightarrow A_1$        &       $10\sim 100$      \\
    $B_2\rightarrow B_1$        &       $10\sim 100$      \\    
    $B_2\rightarrow A_1,\ A_2,\ C$        &       $10\sim 50$      \\    
    $B_1\rightarrow A_1,\ C$        &       $10\sim 500$      \\    \hline\hline
  \end{tabular*}
  \caption{A list of main probabilty fluxes and the corresponding time scales observed in the constant EF case.}
  \label{tab:tmp1}
\end{table}

@@

The Fig.~\ref{fig:tmp3}
shows that the $x$-component of the dipole moment of the alanine dipeptide
reaches 85\% of the maximum value in only 20~ps, which is comparable to
the warm-up time $t_{init}$, and is 25 times
smaller than the slowest time scale of the conformational relaxation.
Then in the following 400~ps, the dipole slowly relaxes to the 
maximum value, that reads
6.8~Debye, and is corresponding to a dipole energy of ca.
$-13.6$~kJ/mol.
The EF induced solvent effect on the alanine is also counted by calculating the
difference of the water-alaine electrostatic interaction 
between systems with and without the EF, and the
energy difference is found to be $-19.1$~~kJ/mol that is
comparable to the dipole energy of the alanine itself.
The main free energy barriers among the
metastable states were reported to be of order $20\sim 40$~kJ/mol~\cite{bohner2012algorithm},
which is comparable to the dipole energy. This explains the 
conformational changes under the electric field.
We calculate the averaged dipole moment of different metastable state
under the constant EF, and they are 6.8, 6.0, 5.1, 3.1 and 7.1~Debye
for $A_1$, $A_2$, $B_1$, $B_2$ and $C$, respectively.
Under a constant EF, the system will be driven towards those
conformations that with higher dipole moment, because the
energy of the system will be lowered by aligning the the dipole
to the external EF.
It is clear that
the $\alpha$-helix and left-handed $\alpha$-helix are of higher
dipole moments than $\beta$-sheet conformations, so the probability
of being in the $\alpha$-helix conformations were
observed being increased under the EF.  $A_1$ and $C$ are the most
probable conformations because their dipole moments are the highest
among all metastable conformation. Moreover, the $B_2$  vanishes, because
its dipole is significantly lower than other conformations.
% This also explains the two-time scale behavior of the molecular dipole
% moment.
From 0 to 20~ps, the probability of
lowest dipole conformation $B_2$ quickly
vanishes, so there is a sharp increment of the average molecular
dipole moment. Then until 500~ps, since the conformation slowly migrates from
$B_1$ to $A_1$ and $C$, and also since the dipole moment $A_1$ and $C$
is higher than $B_1$, the average molecular dipole moment slowly increases
to the maximum value.
